---
title:  "Example Fire Danger Maps for Historical Wet and Dry Periods"
author: 
  - Stephen Huysman, shuysman@gmail.com
email: "shuysman@gmail.com"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    theme: cerulean
    number_sections: true
    toc: true
---

```{r setup}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.height = 12, fig.width = 24)
```

# Example Fire Danger Maps for Historical Wet and Dry Periods

This section generates fire danger maps for one-week periods during notable historical drought years (1988, 2016) and wet years (2011, 1993) to illustrate the model's behavior under different conditions.

First, pre-calculate the rolling sums for the entire time series to avoid re-computing them for each example. This is memory-intensive but more efficient.

```{r}
library(tidyverse)
library(terra)
library(tidyterra)
library(glue)
library(maptiles)

replace_duplicated <- function(x) {
  x[duplicated(x)] <- NA
  return(x)
}

bin_rast <- function(new_rast, quants_rast, probs) {
  # Count how many quantile layers the new value is greater than.
  # This results in a raster of integers from 0 to 9.
  bin_index_rast <- sum(new_rast > quants_rast)

  # Now, map this integer index back to a percentile value.
  # We need a mapping from [0, 1, 2, ..., 9] to [0, 0.1, 0.2, ..., 0.9]
  # A value of 0 means it was smaller than the 1st quantile (q_0.1)
  # A value of 9 means it was larger than the 9th quantile (q_0.9)
  percentile_map <- c(0, probs)
  from_vals <- 0:length(probs)
  rcl_matrix <- cbind(from_vals, percentile_map)

  # Use classify to create the final approximate percentile raster
  percentile_rast_binned <- classify(bin_index_rast, rcl = rcl_matrix)

  return(percentile_rast_binned)
}

terraOptions(
  verbose = FALSE,
  memfrac = 0.9
)

## Optimal rolling windows determined by dryness analysis script

probs <- seq(.01, 1.0, by = .01)

nps_boundaries <- vect("data/nps_boundary/nps_boundary.shp") %>%
  filter(UNIT_CODE %in% c("YELL", "GRTE", "JODR"))

vpd_data_dir <- file.path("/media/steve/THREDDS/gridmet/")
vpd_data_files <- list.files(vpd_data_dir, pattern = "vpd.*.nc", full.names = TRUE)
vpd_data <- rast(vpd_data_files) %>%
  crop(project(nps_boundaries, crs(.))) %>%
  mask(project(nps_boundaries, crs(.)))

time(vpd_data) <- as_date(depth(vpd_data), origin = "1900-01-01")

forest_quants_rast <- rast("./out/ecdf/17-middle_rockies-forest/17-middle_rockies-forest-5-VPD-quants.nc")
non_forest_quants_rast <- rast("./out/ecdf/17-middle_rockies-forest/17-middle_rockies-forest-5-VPD-quants.nc")

cover_types <- rast("data/LF2023_EVT_240_CONUS/Tif/4326/LC23_EVT_240.tif") %>%
  crop(vpd_data)

activeCat(cover_types) <- "EVT_LF"

# 1. Extract the category levels into a data frame
# The [[1]] is used because levels() returns a list, one for each layer.
categories_df <- levels(cover_types)[[1]]

# 2. Add a new column with the desired classification using your rules
# We will create numeric codes: 1 for non_forest, 2 for forest.
# Everything else will become NA.
categories_df <- categories_df %>%
  mutate(veg_class_id = case_match(
    EVT_LF,
    c("Herb", "Shrub", "Sparse") ~ 1,
    "Tree" ~ 2,
    .default = NA_integer_
  ))

# 3. Create the reclassification matrix from the original ID to the new ID
# The matrix should have two columns: 'from' (original ID) and 'to' (new ID).
# The original raster values are in the 'ID' or 'value' column of the levels table.
rcl_matrix <- categories_df[, c("Value", "veg_class_id")]

# 4. Classify the raster using this matrix
classified_rast <- classify(cover_types, rcl = rcl_matrix, right = NA)

# 5. (Optional but recommended) Assign new category labels to the output raster
new_levels <- data.frame(
  ID = c(1, 2),
  cover = c("non_forest", "forest")
)
levels(classified_rast) <- new_levels

forest_mask <- classified_rast == 2
forest_mask <- subst(forest_mask, FALSE, NA)
non_forest_mask <- classified_rast == 1
non_forest_mask <- subst(non_forest_mask, FALSE, NA)

basemap <- get_tiles(classified_rast, provider = "Esri.NatGeoWorldMap", zoom = 9) %>%
  crop(classified_rast)
```


```{r}
forest_test_full <- terra::roll(vpd_data, n = 5, fun = mean, type = "to", circular = FALSE, overwrite = TRUE)
non_forest_test_full <- terra::roll(vpd_data, n = 21, fun = mean, type = "to", circular = FALSE, overwrite = TRUE)
```

## 1988 Drought Example

A one-week period in August 1988, during the height of the Yellowstone fires.

```{r}
test_dates <- as_date("1988-08-08") + 0:6

forest_test <- forest_test_full %>% subset(time(.) %in% test_dates)
non_forest_test <- non_forest_test_full %>% subset(time(.) %in% test_dates)

forest_fire_danger_rast <- rast()
for (n in 1:nlyr(forest_test)) {
  forest_percentile_rast <- bin_rast(subset(forest_test, n), forest_quants_rast, probs)
  forest_fire_danger_ecdf <- readRDS("/home/steve/sync/pyrome-fire/out/ecdf/17-middle_rockies-forest/17-middle_rockies-forest-5-VPD-ecdf.RDS")
  forest_fire_danger_rast <- c(forest_fire_danger_rast, terra::app(forest_percentile_rast, fun = \(x) forest_fire_danger_ecdf(x)))
}

non_forest_fire_danger_rast <- rast()
for (n in 1:nlyr(non_forest_test)) {
  non_forest_percentile_rast <- bin_rast(subset(non_forest_test, n), non_forest_quants_rast, probs)
  non_forest_fire_danger_ecdf <- readRDS("/home/steve/sync/pyrome-fire/out/ecdf/17-middle_rockies-non_forest/17-middle_rockies-non_forest-21-VPD-ecdf.RDS")
  non_forest_fire_danger_rast <- c(non_forest_fire_danger_rast, terra::app(non_forest_percentile_rast, fun = \(x) non_forest_fire_danger_ecdf(x)))
}

forest_fire_danger_rast <- resample(forest_fire_danger_rast, classified_rast)
non_forest_fire_danger_rast <- resample(non_forest_fire_danger_rast, classified_rast)

names(forest_fire_danger_rast) <- test_dates
names(non_forest_fire_danger_rast) <- test_dates

ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(data = mask(forest_fire_danger_rast, forest_mask)) +
  geom_spatraster(data = mask(non_forest_fire_danger_rast, non_forest_mask)) +
  scale_fill_viridis_c(option = "B", na.value = "transparent", limits = c(0, 1)) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "Fire Danger: August 1988 (Drought)", fill = "Fire Danger")
```

## 2016 Drought Example

A one-week period during a mid-August drying spell in 2016.

```{r}
test_dates <- as_date("2016-08-15") + 0:6

forest_test <- forest_test_full %>% subset(time(.) %in% test_dates)
non_forest_test <- non_forest_test_full %>% subset(time(.) %in% test_dates)

forest_fire_danger_rast <- rast()
for (n in 1:nlyr(forest_test)) {
  forest_percentile_rast <- bin_rast(subset(forest_test, n), forest_quants_rast, probs)
  forest_fire_danger_ecdf <- readRDS("/home/steve/sync/pyrome-fire/out/ecdf/17-middle_rockies-forest/17-middle_rockies-forest-5-VPD-ecdf.RDS")
  forest_fire_danger_rast <- c(forest_fire_danger_rast, terra::app(forest_percentile_rast, fun = \(x) forest_fire_danger_ecdf(x)))
}

non_forest_fire_danger_rast <- rast()
for (n in 1:nlyr(non_forest_test)) {
  non_forest_percentile_rast <- bin_rast(subset(non_forest_test, n), non_forest_quants_rast, probs)
  non_forest_fire_danger_ecdf <- readRDS("/home/steve/sync/pyrome-fire/out/ecdf/17-middle_rockies-non_forest/17-middle_rockies-non_forest-21-VPD-ecdf.RDS")
  non_forest_fire_danger_rast <- c(non_forest_fire_danger_rast, terra::app(non_forest_percentile_rast, fun = \(x) non_forest_fire_danger_ecdf(x)))
}

forest_fire_danger_rast <- resample(forest_fire_danger_rast, classified_rast)
non_forest_fire_danger_rast <- resample(non_forest_fire_danger_rast, classified_rast)

names(forest_fire_danger_rast) <- test_dates
names(non_forest_fire_danger_rast) <- test_dates

ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(data = mask(forest_fire_danger_rast, forest_mask)) +
  geom_spatraster(data = mask(non_forest_fire_danger_rast, non_forest_mask)) +
  scale_fill_viridis_c(option = "B", na.value = "transparent", limits = c(0, 1)) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "Fire Danger: Mid-August 2016 (Drought)", fill = "Fire Danger")
```

## 2011 Wet Period Example

A one-week period in July during the historically wet year of 2011.

```{r}
test_dates <- as_date("2011-07-20") + 0:6

forest_test <- forest_test_full %>% subset(time(.) %in% test_dates) 
non_forest_test <- non_forest_test_full %>% subset(time(.) %in% test_dates)

forest_fire_danger_rast <- rast()
for (n in 1:nlyr(forest_test)) {
  forest_percentile_rast <- bin_rast(subset(forest_test, n), forest_quants_rast, probs)
  forest_fire_danger_ecdf <- readRDS("/home/steve/sync/pyrome-fire/out/ecdf/17-middle_rockies-forest/17-middle_rockies-forest-5-VPD-ecdf.RDS")
  forest_fire_danger_rast <- c(forest_fire_danger_rast, terra::app(forest_percentile_rast, fun = \(x) forest_fire_danger_ecdf(x)))
}

non_forest_fire_danger_rast <- rast()
for (n in 1:nlyr(non_forest_test)) {
  non_forest_percentile_rast <- bin_rast(subset(non_forest_test, n), non_forest_quants_rast, probs)
  non_forest_fire_danger_ecdf <- readRDS("/home/steve/sync/pyrome-fire/out/ecdf/17-middle_rockies-non_forest/17-middle_rockies-non_forest-21-VPD-ecdf.RDS")
  non_forest_fire_danger_rast <- c(non_forest_fire_danger_rast, terra::app(non_forest_percentile_rast, fun = \(x) non_forest_fire_danger_ecdf(x)))
}

forest_fire_danger_rast <- resample(forest_fire_danger_rast, classified_rast)
non_forest_fire_danger_rast <- resample(non_forest_fire_danger_rast, classified_rast)

names(forest_fire_danger_rast) <- test_dates
names(non_forest_fire_danger_rast) <- test_dates

ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(data = mask(forest_fire_danger_rast, forest_mask)) +
  geom_spatraster(data = mask(non_forest_fire_danger_rast, non_forest_mask)) +
  scale_fill_viridis_c(option = "B", na.value = "transparent", limits = c(0, 1)) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "Fire Danger: Late July 2011 (Wet)", fill = "Fire Danger")
```

## 1993 Wet Period Example

A one-week period in July during the historically wet year of 1993.

```{r}
test_dates <- as_date("1993-07-20") + 0:6

forest_test <- forest_test_full %>% subset(time(.) %in% test_dates)
non_forest_test <- non_forest_test_full %>% subset(time(.) %in% test_dates)

forest_fire_danger_rast <- rast()
for (n in 1:nlyr(forest_test)) {
  forest_percentile_rast <- bin_rast(subset(forest_test, n), forest_quants_rast, probs)
  forest_fire_danger_ecdf <- readRDS("/home/steve/sync/pyrome-fire/out/ecdf/17-middle_rockies-forest/17-middle_rockies-forest-5-VPD-ecdf.RDS")
  forest_fire_danger_rast <- c(forest_fire_danger_rast, terra::app(forest_percentile_rast, fun = \(x) forest_fire_danger_ecdf(x)))
}

non_forest_fire_danger_rast <- rast()
for (n in 1:nlyr(non_forest_test)) {
  non_forest_percentile_rast <- bin_rast(subset(non_forest_test, n), non_forest_quants_rast, probs)
  non_forest_fire_danger_ecdf <- readRDS("/home/steve/sync/pyrome-fire/out/ecdf/17-middle_rockies-non_forest/17-middle_rockies-non_forest-21-VPD-ecdf.RDS")
  non_forest_fire_danger_rast <- c(non_forest_fire_danger_rast, terra::app(non_forest_percentile_rast, fun = \(x) non_forest_fire_danger_ecdf(x)))
}

forest_fire_danger_rast <- resample(forest_fire_danger_rast, classified_rast)
non_forest_fire_danger_rast <- resample(non_forest_fire_danger_rast, classified_rast)

names(forest_fire_danger_rast) <- test_dates
names(non_forest_fire_danger_rast) <- test_dates

ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(data = mask(forest_fire_danger_rast, forest_mask)) +
  geom_spatraster(data = mask(non_forest_fire_danger_rast, non_forest_mask)) +
  scale_fill_viridis_c(option = "B", na.value = "transparent", limits = c(0, 1)) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "Fire Danger: Late July 1993 (Wet)", fill = "Fire Danger")
```
